% Anonymous submission toggle:
% - \anonsubmissiontrue  => anonymous submission build
% - \anonsubmissionfalse => named draft / camera-ready build
\newif\ifanonsubmission
\anonsubmissiontrue
\ifanonsubmission
  \PassOptionsToClass{anonymous}{acmart}
\fi
\documentclass[sigconf]{acmart}

\usepackage{amsmath}
\usepackage{booktabs}

\setcopyright{none}
\settopmatter{printfolios=true}
\acmDOI{}
\acmISBN{}
\acmConference[NSPW '26]{New Security Paradigms Workshop}{2026}{TBD}
\acmBooktitle{Proceedings of the New Security Paradigms Workshop (NSPW '26)}

\title{Hardware Security is Security Economics:\\ A Unifying Synthesis and Ontology}

\author{Author Name}
\affiliation{%
  \institution{Affiliation}
  \city{City}
  \country{Country}
}
\email{email@example.com}

\begin{document}
\begin{abstract}
Every security mechanism hides a distributional question that its designers rarely ask
explicitly: \emph{who pays?} The performance tax from speculative execution mitigations
fell on cloud operators who had no role in the original design decision. The integration
burden from capability-based memory safety falls on software ecosystems that must
retrofit decades of code. The compliance overhead from trusted boot requirements falls
on OEMs navigating certification timelines they did not set. In each case, the technical
choice and its economic consequence are made by different actors, recorded in different
documents, and analyzed by different communities---if they are analyzed at all.

This paper argues that this disconnect is not accidental. It is the product of two
traditions that have been analyzing the same underlying problem with incompatible
vocabularies: systems engineering, which characterizes the technical tradeoff space with
precision but treats cost distribution as out of scope, and security economics, which
understands distribution and incentive misalignment clearly but treats the mechanism
as a black box. Neither tradition, alone, can explain why a given mechanism produces
the burden distribution it does---or what alternative designs would have implied a
different one. We propose a synthesis: a burden-allocation framework in which security
mechanisms are understood as instruments for distributing costs, risks, and constraints
across stakeholders and timescales. A lightweight shared vocabulary, grounded in two
case studies, makes the translation between engineering and economic concepts precise
enough to be useful in practice.
\end{abstract}

\maketitle

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10003022.10003023</concept_id>
<concept_desc>Security and privacy~Embedded systems security</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002978.10003022.10003036</concept_id>
<concept_desc>Security and privacy~Systems security</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[500]{Security and privacy~Embedded systems security}
\ccsdesc[300]{Security and privacy~Systems security}

\keywords{hardware security, ontology, security economics, computer architecture, burden-allocation instruments, risk transfer}

\input{sections/introduction}
\input{sections/twotraditions}
\input{sections/casestudy1}
\input{sections/casestudy2}
\input{sections/vocabulary}
\input{sections/usage}
\input{sections/limitations}


% \input{sections/paradigm}
% \input{sections/background}
% \input{sections/ontology}
% \input{sections/analytical-framework}
% \input{sections/case-studies}
% \input{sections/methodology}
% \input{sections/evaluation}
% \input{sections/discussion}
\input{sections/conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}
